State of the Practice for Lattice Boltzmann Method Software

### 2022-01-25 StateOfPractice/Papers/LatticeBoltz_SOP-arXiv/LatticeBoltz_SOP.tex
Edits up to and including Section 3.1 on Installability.

Answers \rqref{RQ_HighestQuality}.  Before discussing quality differences between the packages and the results of
the AHP ranking, we will contrast the different packages based on features, as
shown in Table~\ref{tbl_features}. \wss{Add some discussion of some example
features.  Add Non-Newtonian and Fluid Structure interaction.  Listed in
alphabetical order.}  The sections below summarize how the software packages
compare based on qualities, as measured by the measurement template.   The
results of the AHP ranking are also discussed.

@@

@@

All of the 24 software packages that were tested have installation instructions.
As noted previously, many of the 22 software packages that were part of the
original long list of 46 packages were removed due to not including
documentation or installation instructions. Of the 24 software packages that
were tested, most had installation instructions located in one place, often in
an instruction manual or on a web-page. Sometimes, like with Ludwig, incomplete
installation instructions are found on a home page, with more detailed
instructions located on another web-page, or within the documentation.
Maintainability and correctness of these instructions could be improved if all
the instructions were in one location.

All 24 packages on the short list can be installed on some Unix-like systems.
Eight packages could be installed on Windows, and five on macOS. Operating
system requirements are listed in the documentation of 20 software packages, but
compatible operating system versions are only specified in four packages
(ESPResSo, LUMA, Mechsys, Sailfish). All but one of the software packages (TCLB)
were tested on Ubuntu for this state of the practice assessment. TCLB was tested
on CentOS, since this operating system is mentioned in its installation
instructions.

@@

Errors encountered during the installation process were often quickly fixed
thanks to descriptive error messages. Systems that provided vague error
messages, such as messages that did not specify which action or file was at
fault, were more difficult to troubleshoot. Only three software packages
(HemeLB, LB3D, lbmpy) that displayed a descriptive error message were not
recoverable, and most of these instances were due to hardware and operating
system incompatibility, such as the requirement of CUDA. Fourteen software
packages definitively broke during installation. Some packages, such as
LB2D-Prime and LB3D-Prime, did not provide a definitive message of the success
or failure of installation. In these instances, validating the installation
required performing a tutorial or running a script, as described below, if these
were available.

@@

Seventeen software packages require less than 10 dependencies to be installed. All but one software package (LatBo.jl) require less than 20 dependencies. Some
packages may automatically install additional dependencies in the background.
Twenty of the software packages do not explicitly indicate software dependency
versions. Some software package installation issues, specifically those
occurring when manual installation of dependencies is required, may be avoided
if versions of dependencies are specified. Sixteen software packages do not have
detailed instructions for installing dependencies. Sixteen software packages
have less than 10 manual installation steps. If dependencies are installed in
one command then none of the software packages take more than 20 steps to
install. The average number of steps is about eight, and the fewest is two
(LB3D-Prime).

@@

Figure~\ref{Fig_Installability} shows the installability ranking of the software
packages using AHP. Software packages with a higher score (ESPResSo, MechSys, OpenLB, DL\_MESO, Palabos) tend to have one set of linear installation instructions that are written as if the person doing the installation has none of the dependencies installed. The instructions often list compatible operating system versions and include instructions for the installation of dependencies. The top ranked packages often incorporate some sort of automation of the installation process and have fewer manual installation steps. The number of dependencies a package has does not correlate with a higher score. The ability to validate the installation process, often through tutorials or test examples that include expected output, is correlated with a higher score. Furthermore, the top seven ranked packages are noted as being alive.

Many software packages would benefit from a rewrite or reorganization of
installation instructions. A single location for installation instructions would
improve their maintainability and correctness. Listing compatible operating
system and dependency versions would decrease installation time and errors, as
would adding instructions on installing dependencies. Installation process
errors should prompt the system to display detailed messages. Once a software
package is installed, either an automatic validation needs to be performed or
the user needs to be able to perform a manual validation using test examples
that include expected output. Finally, uninstallation instructions should be
included in the documentation.

@@

\subsection{Reproducibility}

@@

Answers \rqref{RQ_Recommend}.  Recommendations for Addressing Developer
Concerns.  This section answers the research questions listed in
Section~\ref{secIntro} using the quantitative data presented in
Section~\ref{AHPresults}, qualitative data presented in
Section~\ref{interviewresults}, additional data from software package
repositories and artifacts, and domain expert feedback.

@@

Regarding data collection, a risk to the validity of this assessment is missing
or incorrect data. Some software package data may not have been measured due to
technology issues like broken links. This issue arose with the measurement of
Palabos, which had a broken link to its user manual, as noted in
Section~\ref{artifacts}.

@@

@@

Section~\ref{interviewresults} noted that developers were not asked to comment
on the reproducibility of their source code from their requirements
specifications and design documentation. Adding this question to the interview
guide should be considered in the next iteration of state of the practice
assessments. Furthermore, as mentioned in Section~\ref{threats}, it was found
that some pertinent information was not specified in public artifacts. The use
of unit testing and continuous integration by several packages (OpenLB, pyLBM,
TCLB) was only discovered during interviews with developers. Adding further
questions to the interview guide regarding the measures that are on the
measurement template could reduce instances of incorrect data being collected.
This additional interview data could be analyzed and incorporated into the AHP
ranking, ensuring quality designations more accurate represent the true quality
of the software packages.  \wss{clarify the distinction between usability and
understandability.}

>>>>

In this section we answer \rqref{RQ_HighestQuality} by using AHP to rank the
identified LBM software against best practices.  Before discussing quality
differences between the packages and the results of the AHP ranking, we will
contrast the different packages based on their features, as shown in
Table~\ref{tbl_features}. The set of LBM programs form a program family, or a
software product line.  The packages all have the commonality of using LBM
techniques to simulate the motion of fluids, but there are also variabilities
between them.  LBM packages can be distinguished by the dimension of problems
they simulate, whether or not (and how) they support parallel computation,
whether or not compressibility can be considered, whether reflexive boundaries
can be employed, whether the simulation handles multi-fluids, whether the
simulations handles turbulence, and whether complex geometries can be simulated.
All of these variabilities are listed as columns in Table~\ref{tbl_features}
with the software listed in alphabetical order. A more complete analysis of the
commonalities and variabilities between LBM packages is provided in
\citet{Michalski2021}.

@@

The sections below summarize how the software packages compare based on the
qualities discussed in Section~\ref{softwarequalities}, as measured via the
measurement template (Section~\ref{empiricalmeasures}).  For the purpose of
discussion, the last subsection combines all the qualities into one ``global''
ranking.

@@

All of the 24 software packages that were tested have installation instructions.
Most have their installation instructions located in one place, often in an
instruction manual or on a web-page. Sometimes, like with Ludwig, incomplete
installation instructions are found on a home page, with more detailed
instructions located on another web-page, or within the documentation.
Installability of the software, and maintainability and correctness of the
instructions themselves, are improved if all the instructions are in one
location.

All 24 packages on the short list can be installed on a Unix-like system. Eight
packages could be installed on Windows, and five on macOS. Operating system
requirements are listed in the documentation of 20 software packages, but
compatible operating system versions are only specified in four packages
(ESPResSo, LUMA, Mechsys, Sailfish). All but one of the software packages (TCLB)
were tested on Ubuntu for this state of the practice assessment. TCLB was tested
on CentOS, since CentOS is mentioned in its installation instructions.

@@

Errors encountered during the installation process were often quickly fixed
thanks to descriptive error messages. Systems that provided vague error
messages, such as messages that did not specify which action or file was at
fault, were more difficult to troubleshoot. Only three software packages
(HemeLB, LB3D, lbmpy) that displayed a descriptive error message were not
recoverable, and most of these instances were due to hardware and operating
system incompatibility, such as the requirement of CUDA. Fourteen software
packages broke during installation. Some packages, such as LB2D-Prime and
LB3D-Prime, did not provide a definitive message of the success or failure of
installation. In these instances, validating the installation required
performing a tutorial or running a script, as described below, if these were
available.

@@

Seventeen software packages require less than 10 dependencies to be installed.
All but one software package (LatBo.jl) require less than 20 dependencies. Some
packages may automatically install additional dependencies in the background.
Twenty of the software packages do not explicitly indicate the versions of
software dependencies. Some software package installation issues, specifically
those occurring when manual installation of dependencies is required, may be
avoided if versions of dependencies are specified. Sixteen software packages do
not have detailed instructions for installing dependencies. Sixteen software
packages have less than 10 manual installation steps. If dependencies are
installed in one command then none of the software packages take more than 20
steps to install. The average number of steps is about eight, and the fewest is
two (LB3D-Prime).

@@

Figure~\ref{Fig_Installability} shows the installability ranking of the software
packages using AHP. Software packages with a higher score (ESPResSo, MechSys,
OpenLB, DL\_MESO, Palabos) tend to have one set of linear installation
instructions written as if the person doing the installation has none of the
dependencies installed. The instructions often list compatible operating system
versions and include instructions for the installation of dependencies. The top
ranked packages often incorporate some sort of automation of the installation
process and have fewer manual installation steps. The number of dependencies a
package has does not correlate with a higher score. The ability to validate the
installation process, often through tutorials or test examples that include
expected output, is correlated with a higher score. Active work on a project
seems to improve installability, since the top seven ranked packages are all
classified as alive.

@@

\subsection{Reproducibility} \label{SecReproducibility}

@@

Answers \rqref{RQ_Recommend}.  Recommendations for Addressing Developer
Concerns.  This section answers the research questions listed in
Section~\ref{secIntro} using the quantitative data presented in
Section~\ref{AHPresults}, qualitative data from developer interviews
(Section~\ref{SecSurvey}), additional data from software package repositories
and artifacts, and domain expert feedback.

@@

Regarding data collection, a risk to the validity of this assessment is missing
or incorrect data. Some software package data may not have been measured due to
technology issues like broken links. This issue arose with the measurement of
Palabos, which had a broken link to its user manual, as noted in
Section~\ref{Sec_CompareArtifacts}.

@@

Many software packages would benefit from a rewrite or reorganization of
installation instructions. A single location for installation instructions would
improve their maintainability and correctness. Listing compatible operating
system and dependency versions would decrease installation time and errors, as
would adding instructions on installing dependencies. Installation process
errors should prompt the system to display detailed messages. Once a software
package is installed, either an automatic validation needs to be performed or
the user needs to be able to perform a manual validation using test examples
that include expected output. Finally, uninstallation instructions should be
included in the documentation. \wss{Have to decide how much to emphasizse
installability.}  \wss{Maybe at some point should mention more modern
installation support, since most of the projects use make, and there are other
options.}

@@

Section~\ref{SecReproducibility} noted that developers were not asked to comment
on the reproducibility of their source code from their requirements
specifications and design documentation. Adding this question to the interview
guide should be considered in the next iteration of state of the practice
assessments. Furthermore, as mentioned in Section~\ref{threats}, it was found
that some pertinent information was not specified in public artifacts. The use
of unit testing and continuous integration by several packages (OpenLB, pyLBM,
TCLB) was only discovered during interviews with developers. Adding further
questions to the interview guide regarding the measures that are on the
measurement template could reduce instances of incorrect data being collected.
This additional interview data could be analyzed and incorporated into the AHP
ranking, ensuring quality designations more accurate represent the true quality
of the software packages.  \wss{clarify the distinction between usability and
understandability.}
